---
title: "BM1 final project"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include = FALSE}
library(tidyverse)
rm(list = ls())
library(arsenal)
library(ggplot2)
library(HH)
library(leaps)
library(jtools)


theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r, echo = F, message = FALSE, warning=FALSE}
# Data input
crimes<-read.csv("HateCrimes.csv") %>% 
  drop_na()
# Data format transformation
crimes = 
  crimes %>% 
  mutate(
    unemployment = as.factor(unemployment),
    unemployment = fct_relevel(unemployment, "low", "high"),
    urbanization = as.factor(urbanization),
    urbanization = fct_relevel(urbanization, "low", "high"),
    hate_crimes_per_100k_splc = as.numeric(crimes$hate_crimes_per_100k_splc))
str(crimes)
```

## Descriptive statistics+Graphs
```{r, echo = F, message = FALSE, warning=FALSE}
# Descriptive Characteristics of study sample
labels1 <- list(hate_crimes_per_100k_splc = "Hate crime rate per 100k population", unemployment = "Level of state unemployment", urbanization="Level of state urbanization", median_household_income="Median household income per state",perc_population_with_high_school_degree="Percentage of adults (>25 yrs) with a high school degree",perc_non_citizen="Percentage of population that are not US citizens",gini_index="Gini index",perc_non_white="Percentage of population that are non-white")

controls1 <- tableby.control(total = T,test=F, numeric.stats = c("meansd", "median", "iqr","range", "Nmiss2"),cat.stats = c("countpct", "Nmiss2"),stats.labels = list(meansd = "Mean/SD",median = "Median",iqr="IQR",range = "Min - Max",Nmiss2 = "Missing",countpct = "N (%)"),digits = 2L,digits.count = 0L)
tb1<-tableby(~hate_crimes_per_100k_splc+unemployment+urbanization+median_household_income+perc_population_with_high_school_degree+perc_non_citizen+gini_index+perc_non_white,data=crimes,control = controls1)
summary(tb1, title = "Descriptive Characteristics of Study Sample, Hate Crimes Occurring in the United States (By State)", labelTranslations = labels1, text=T)

# Histogram of the outcome distribution
mytheme <- theme_bw() + theme(plot.title=element_text(hjust=0.5),axis.title=element_text(size=rel(1)),axis.text=element_text(size=rel(1)),legend.position = "top")
ggplot(data=crimes,aes(x=hate_crimes_per_100k_splc))+
geom_histogram(fill="lightblue",color="darkblue",linetype="dashed",alpha=0.5) +
labs(title="Histogram of Hate Crimes Rate Per 100k Population in United States",x="Hate Crimes Rate (Per 100 K Population)", y = "Count")+mytheme
## District of columbia might be a outlier

# 2. Boxplots for each variable to see if there is any outlier or influential points.
par(mfrow=c(2,3))
boxplot(crimes$hate_crimes_per_100k_splc, main='Hate Crimes Rate')
boxplot(crimes$median_household_income, main='Median Household Income')
boxplot(crimes$perc_population_with_high_school_degree,main='Percentage with Highschool Degree' )
boxplot(crimes$perc_non_citizen, main='Percentage of Non-Citizen')
boxplot(crimes$perc_non_white, main='Percentage of Non-White')
boxplot(crimes$gini_index, main='Gini Index')
## District of columbia might be a outlier
```

## Varible Selection
```{r, echo = F, message = FALSE, warning=FALSE}
# Correlation matrix
library("corrplot")
matrix1<-cor(crimes[,-c(1:3)],use="complete.obs")
matrix1 %>% 
    corrplot::corrplot(
    method = "square",
    type = "lower",
    addCoef.col = "black", 
    diag = T,
    tl.cex = 0.5
  )
round(matrix1,2) # Cor(non-white and non-citizen): 0.75; cor(income and > high school degree):0.65

```

median_household_income and perc_population_with_high_school_degree may highly correlated.
perc_non_white and perc_non_cizen may highly correlated.

```{r}
# 2. Simple linear regression model: hate crime rates~ gini index
ggplot(data=crimes,aes(x=gini_index,y=hate_crimes_per_100k_splc))+
geom_point(color="blue",alpha=0.5) +geom_smooth(method="lm",se=F,color="pink")+labs(title="Plot of Hate Crimes Rate By Gini Index",y="Hate Crimes Rate (Per 100 K Population)", x = "Gini Index")+mytheme
res_crimes_gini<-lm(hate_crimes_per_100k_splc~gini_index,data=crimes)
summary(res_crimes_gini)

# 3. Varibale Selection- backwards selection
res1 <- lm(hate_crimes_per_100k_splc ~ .-state, data=crimes)
summary(res1)

step1<-update(res1, . ~ . -perc_non_white)
summary(step1)

step2<-update(step1, . ~ . -unemployment)
summary(step2)

step3<-update(step2, . ~ . -median_household_income) 
summary(step3)

step4<-update(step3, . ~ . -perc_non_citizen) 
summary(step4)

step5<-update(step4, . ~ . -urbanization)
summary(step5)

multi_fit1<-lm(hate_crimes_per_100k_splc~perc_population_with_high_school_degree+gini_index,data=crimes)
summary(multi_fit1)

# 4. Criterion Selection
best <- function(model, ...) 
{
  subsets <- regsubsets(formula(model), model.frame(model), ...)
  subsets <- with(summary(subsets),
                  cbind(p = as.numeric(rownames(which)), which, rss, rsq, adjr2, cp, bic))
  
  return(subsets)
}  


round(best(multi_fit1, nbest = 1), 3) # chose model2 with the minimum cp, BIC and highest adjusted R2

```

### just one step to get best fit.
```{r}
crimes = 
  crimes %>% 
  subset(select = c(-state))

model_fit = 
  lm(hate_crimes_per_100k_splc ~., data = crimes) %>% 
  step(direction = "both")
summary(model_fit)   ### get same model with Mumei
```

### interaction

```{r}
# 5. Check interactions between median income and perc of high school degree
res_interaction<-lm(hate_crimes_per_100k_splc ~ gini_index + perc_population_with_high_school_degree * median_household_income ,data=crimes)
summary(res_interaction) # No-interaction between income and high school degree
anova(model_fit, res_interaction)

#--------------------------------------------------------------------------------------------------------

# Check interactions between urbanization and gini_index
ggplot(data=crimes,aes(x=gini_index,y=hate_crimes_per_100k_splc,color=urbanization))+
geom_point() +geom_smooth(method="lm",se=F)+labs(title="Plot of Hate Crimes Rate vs Gini index By Urbanization level ",y="Hate Crimes Rate (Per 100 K Population)", x = "Gini index")+mytheme



res_interaction2<-lm(hate_crimes_per_100k_splc~perc_population_with_high_school_degree+gini_index*urbanization,data=crimes)
summary(res_interaction2) # No interaction b/w urbanization and gini index

anova(model_fit, res_interaction2)

###  Check interactions between perc_non_white and gini_index

res_interaction3<-lm(hate_crimes_per_100k_splc~perc_population_with_high_school_degree+gini_index*perc_non_white,data=crimes)
summary(res_interaction3) 

anova(model_fit, res_interaction3)#exist interaction b/w perc_non_white and gini index

###  Check interactions between high school degree and unemployment

res_interaction4<-lm(hate_crimes_per_100k_splc~perc_population_with_high_school_degree*unemployment+gini_index,data=crimes)
summary(res_interaction4) 

anova(model_fit, res_interaction4)

###  Check interactions between gini index and unemployment

res_interaction4<-lm(hate_crimes_per_100k_splc~perc_population_with_high_school_degree + unemployment*gini_index,data=crimes)
summary(res_interaction4) 

anova(model_fit, res_interaction4)

# Check interactions between urbanization and high school degree

res_interaction5<-lm(hate_crimes_per_100k_splc~perc_population_with_high_school_degree*urbanization+gini_index,data=crimes)
summary(res_interaction5)
anova(model_fit, res_interaction5)


```
The anova p-value of potential interaction between urbanization and gini index,and median income and perc of high school degree turn out to not be statistically significant. So even though the estimated slopes in the interaction model look very different, our estimates are quite variable, so we don’t have enough evidence to conclude that the interaction term (different slopes) is providing significant additional explanatory power over the simpler perc_population_with_high_school_degree+gini_index. (do not just copy this sentence, paraphrase it)

There is an significant interaction between perc_non_white and gini index



### find confounder

```{r}
reg_conf1 = lm(hate_crimes_per_100k_splc~perc_population_with_high_school_degree+gini_index + unemployment, data=crimes)
summary(model_fit)
summary(reg_conf1) ### coefficient changes for gini and high school degree is smaller than 1%, unemployment is not a confounder

reg_conf2 = lm(hate_crimes_per_100k_splc~perc_population_with_high_school_degree+gini_index + urbanization, data=crimes)
summary(model_fit)
summary(reg_conf2) ### coefficient changes for gini and high school degree is smaller than 1%, urbanization is not a confounder

reg_conf3 = lm(hate_crimes_per_100k_splc~perc_population_with_high_school_degree+gini_index + median_household_income, data=crimes)
summary(model_fit)
summary(reg_conf3)

reg_conf4 = lm(hate_crimes_per_100k_splc~perc_population_with_high_school_degree+gini_index + perc_non_citizen, data=crimes)
summary(model_fit)
summary(reg_conf4) # 5% coefficient change for gini_index

reg_conf5 = lm(hate_crimes_per_100k_splc~perc_population_with_high_school_degree+gini_index + perc_non_white, data=crimes)
summary(model_fit)
summary(reg_conf5)
```

coefficient changes for gini and high school degree is smaller than 1%, none of other variables are confounders

## Model Diagnosis
### check assumption， final model(finish variables selection)
```{r}
final_fit = lm(hate_crimes_per_100k_splc ~ perc_population_with_high_school_degree + gini_index,data=crimes)
summary(final_fit)
par(mfrow=c(2,2))
plot(final_fit) # need transformation
```


## model diagonsis
### box cox

```{r}
boxcox(final_fit)  # use natural log transformation
```

lamda =0, get maximum likelihood, so the outcome should be applied to log transformation.

```{r}
crime_log = 
  crimes %>%
  mutate(ln_hate_crime_outcome = log(hate_crimes_per_100k_splc, base = exp(1))) 
  
final_fit2 = 
  lm(ln_hate_crime_outcome ~ perc_population_with_high_school_degree + gini_index , data = crime_log)
summary(final_fit2)

```

### Identify Collinearity
```{r}
# Calculate the variance inflation factor (VIF)
vif(final_fit2)
```
no VIF higher than 5, nothing should be removed because of collinearity

### check for outliers

```{r}

# Measures of influence:
# Gives DFFITS, Cook's Distance, Hat diagonal elements, and others.

influence.measures(final_fit2)

# Look at the Cook's distance lines and notice obs 36 and 29 as potential Y outliers/influential points

par(mfrow=c(2,2))
plot(final_fit2)


# Remove observations 9
crime_log_rm_9<-crime_log[c(-9),]
mult.fit_no9<- lm(ln_hate_crime_outcome ~ perc_population_with_high_school_degree + gini_index , data = crime_log_rm_9)

summary(mult.fit_no9)

mult.fit_no9 %>% 
  broom::tidy() %>% 
  knitr::kable()

influence.measures(mult.fit_no9)

par(mfrow=c(2,2))
plot(mult.fit_no9)
```





